Foregrounds in Wide-Field Redshifted 21~cm Power Spectra

### 2015-03-16 ms.tex
Make ApJ Proof corrections
from 95314e1:
============

In ms.tex,

Make ApJ Proof corrections.

(*) Change HI to H I
(*) Change double hyphens to single hyphens
(*) Change non-linear to nonlinear
(*) Minor changes
(*) Remove hyphens in line-of-sight
(*) Change "dot-dashed" to "dotted-dashed"
(*) Change notation of fractional degrees
(*) Change itemizd list to enumerated list
(*) Change hour/hours to hr
(*) Remove hyphens in signal-to-noise
(*) Change "towards" to "toward"

In eor.bib,

(*) Update bibtex reference to Taylor et al.(1999)

Add proof1.txt containing responses to proof corrections

\title{Foregrounds in Wide--Field Redshifted 21~cm Power Spectra}

@@

Detection of 21~cm emission of H{\sc i} from the epoch of reionization, at redshifts $z>6$, is limited primarily by foreground emission. We investigate the signatures of wide--field measurements and an all--sky foreground model using the delay spectrum technique that maps the measurements to foreground object locations through signal delays between antenna pairs. We demonstrate interferometric measurements are inherently sensitive to all scales, including the largest angular scales, owing to the nature of wide--field measurements. These wide--field effects are generic to all observations but antenna shapes impact their amplitudes substantially. A dish--shaped antenna yields the most desirable features from a foreground contamination viewpoint, relative to a dipole or a phased array. Comparing data from recent Murchison Widefield Array observations, we demonstrate that the foreground signatures that have the largest impact on the H{\sc i} signal arise from power received far away from the primary field of view. We identify diffuse emission near the horizon as a significant contributing factor, even on wide antenna spacings that usually represent structures on small scales. For signals entering through the primary field of view, compact emission dominates the foreground contamination. These two mechanisms imprint a characteristic {\it pitchfork} signature on the ``foreground wedge'' in Fourier delay space. Based on these results, we propose that selective down--weighting of data based on antenna spacing and time can mitigate foreground contamination substantially by a factor $\sim 100$ with negligible loss of sensitivity.

@@

At the end of the recombination epoch, the Universe was completely neutral. This period, referred to as the {\it Dark Ages} in the Universe's history, is characterized by the localized accumulation of matter under the influence of gravity. And it ended with the formation of the first stars and galaxies which started emitting ultra--violet and X--ray radiation, thereby reionizing the neutral medium in their surroundings. This commenced the epoch of reionization (EoR) --- a period of non--linear growth of matter density perturbations and astrophysical evolution.

Observing redshifted 21~cm radiation generated by the spin flip transition of H{\sc i} has been identified as a direct probe of the EoR \citep{sun72,sco90,mad97,toz00,ili02}. Detecting this signal has recently emerged as a very promising experiment to fill the gaps in our understanding of the Universe's history.

Sensitive instruments such as the Square Kilometre Array (SKA) are required for direct observation and tomography of redshifted H{\sc i}. Numerous pathfinders and precursors to the SKA such as the Murchison Widefield Array \citep[MWA;][]{lon09,tin13,bow13}, the Low Frequency Array \citep[LOFAR;][]{van13}, and the Precision Array for Probing the Epoch of Reionization \citep[PAPER;][]{par10} have become operational with enough sensitivity for a statistical detection of the EoR H{\sc i} power spectrum \citep{bow06,par12a,bea13,dil13,thy13,pob14}. The Hydrogen Epoch of Reionization Array\footnote{\url{http://reionization.org/}} (HERA) is currently under construction using new insights gained with the MWA and PAPER.

A key challenge in the statistical detection of the redshifted H{\sc i} 21~cm signal, via the spatial power spectrum of temperature fluctuations, arises from the contamination by Galactic and extragalactic foregrounds \citep[see, e.g.,][]{dim02,zal04,fur06,ali08,ber09,ber10,gho12}. \citet{mor04} show that the inherent isotropy and symmetry of the EoR signal in frequency and spatial wavenumber ($k$) space make it distinguishable from sources of contamination which are isolated to certain $k$ modes by virtue of their inherent spectral smoothness \citep{mor06,bow09,liu11,par12b,dil13,pob13}. Since this contamination is expected to be several orders of magnitude stronger than the underlying EoR H{\sc i} signal, it is critical to characterize foregrounds precisely in order to reduce their impact on EoR H{\sc i} power spectrum detection sensitivity.

Considerable effort is being made towards understanding the $k$--space behavior of foreground signatures in the observed power spectrum and formulating robust estimators of the true power spectrum \citep{bow09,liu09,dat10,liu11,mor12,tro12,pob13,thy13,dil14,liu14a,liu14b}. A model that provides a generic explanation for the observed foreground power spectrum has emerged, whereby the wide--field (and chromatic) response of the instrument causes the power in smooth spectrum foregrounds to occupy higher $k$--modes into the so-called `wedge'. The conservative foreground strategy, referred to as {\it avoidance}, that has developed alongside this work is to discard $k$--modes which could be contaminated \citep[e.g.,][]{par14}. The more aggressive alternative is to subtract a sky model and regain access to modes that would be discarded by {\it avoidance}. In both cases, which parts of the sky are most critical to either {\it avoid} or {\it subtract} has remained largely uncertain. Here, we focus primarily on extending the {\it avoidance} strategy by identifying foreground components at greatest risk to ``leak'' from foreground modes to EoR modes and proposing a scheme for down--weighting these components.

Foregrounds with intrinsic deviations from spectral smoothness, instruments with high chromaticity, polarization leakage, calibration errors, or approximations in power spectrum analyses can contaminate the true EoR H{\sc i} power spectrum. Here we use existing catalogs and a high fidelity instrumental model to capture both foreground and instrumental chromaticity. To decouple these effects from possible analysis effects, such as those pointed out by \citet{haz13}, we compute power spectra using a {\it per-baseline} approach of \citet{par12b}. This approximates the power spectrum as the inverse Fourier transform of the spectra generated by the instrument's correlator.

In \S\ref{sec:delay-spectrum} we provide an overview of the delay spectrum technique. We investigate signatures generic to all wide--field measurements of EoR power spectra in \S\ref{sec:wide-field}. In \S\ref{sec:sim}, we present the foreground model and a variety of instrument models to rank antenna shapes based on foreground contamination. In \S\ref{sec:MWA}, we describe the MWA setup, summarize the observing parameters, and present the resulting data. Simulations using these observing parameters are compared with the data and analyzed for foreground signatures. We report two important findings: foregrounds that most severely obscure the redshifted 21~cm power spectrum are not caused by emission in the central field of view, but rather by bright objects from near the horizon; and, diffuse Galactic emission plays a significant role hitherto unpredicted. In \S\ref{sec:fg-grading}, we offer an initial description of a more precise foreground avoidance technique that minimizes foreground contamination using prior knowledge of the sky to down--weight adversely contaminated baselines. We present a summary of our work and findings in \S\ref{sec:summary}.

@@

Interferometer array data known as {\it visibilities}, $V_b(f)$, represent correlations between time--series of electric fields measured by different antenna pairs with separation vectors $\boldsymbol{b}$ and then Fourier transformed along the time axis to obtain a spectrum along the frequency ($f$) axis. If $I(\hat{\boldsymbol{s}},f)$ and $A(\hat{\boldsymbol{s}},f)$ are the sky brightness and antenna's directional power pattern, respectively, at different frequencies as a function of direction on the sky denoted by the unit vector ($\hat{\boldsymbol{s}}$), and $W_\textrm{i}(f)$ denotes instrumental bandpass weights, then $V_b(f)$ can be written as:
\begin{align}\label{eqn:obsvis}
  V_b(f) &= \iint\limits_\textrm{sky} A(\hat{\boldsymbol{s}},f)\,I(\hat{\boldsymbol{s}},f)\,W_\textrm{i}(f)\,e^{-i2\pi f\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}}\,\dif\Omega,
\end{align}
where, $c$ is the speed of light, and $\dif\Omega$ is the solid angle element to which $\hat{\boldsymbol{s}}$ is the unit normal vector. This equation is valid in general, including wide--field measurements, and is a slight adaptation from \citet{van34}, \citet{zer38}, and \citet{tho01}.

The {\it delay spectrum}, $\tilde{V}_b(\tau)$, is defined as the inverse Fourier transform of $V_b(f)$ along the frequency coordinate:
\begin{align}\label{eqn:delay-transform}
  \tilde{V}_b(\tau) &\equiv \int V_b(f)\,W(f)\,e^{i2\pi f\tau}\,\dif f,
\end{align}
where, $W(f)$ is a spectral weighting function which can be chosen to control the quality of the delay spectrum \citep{thy13,ved12}, and $\tau$ represents the signal delay between antenna pairs:
\begin{equation}\label{eqn:delay}
  \tau = \frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}.
\end{equation}
The delay transform conventions used in this paper are described in appendix~\ref{sec:delay-conventions}. $\tilde{V}_b(\tau)$ is expressed in observer's units of Jy~Hz.

The delay spectrum has a close resemblance to cosmological H{\sc i} spatial power spectrum. Appendix~\ref{sec:EoR-power-spectrum} gives an overview of the similarities and differences between the two. Foregrounds can be described in either framework. For our study, we find the delay spectrum approach to be simple and yet extremely useful.

In order to express a quantity derived from $\tilde{V}_b(\tau)$ whose units are the same as that of the cosmological H{\sc i} power spectrum, we define the delay power spectrum:
\begin{align}\label{eqn:delay-power-spectrum}
  P_\textrm{d}(\boldsymbol{k}_\perp,k_\parallel) &\equiv |\tilde{V}_b(\tau)|^2\left(\frac{A_\textrm{e}}{\lambda^2\Delta B}\right)\left(\frac{D^2\Delta D}{\Delta B}\right)\left(\frac{\lambda^2}{2k_\textrm{B}}\right)^2,
\end{align}
with
\begin{align}
  \boldsymbol{k}_\perp &\equiv \frac{2\pi(\frac{\boldsymbol{b}}{\lambda})}{D}, \\
  k_\parallel &\equiv \frac{2\pi\tau\,f_{21}H_0\,E(z)}{c(1+z)^2}, 
\end{align}
where, $A_\textrm{e}$ is the effective area of the antenna, $\Delta B$ is the bandwidth, $\lambda$ is the wavelength of the band center, $k_\textrm{B}$ is the Boltzmann constant, $f_{21}$ is the rest frame frequency of the 21~cm spin flip transition of H{\sc i}, $z$ is the redshift, $D\equiv D(z)$ is the transverse comoving distance, $\Delta D$ is the comoving depth along the line of sight corresponding to $\Delta B$, and $h$, $H_0$ and $E(z)\equiv [\Omega_\textrm{M}(1+z)^3+\Omega_\textrm{k}(1+z)^2+\Omega_\Lambda]^{1/2}$ are standard terms in cosmology. Throughout the paper, we use $\Omega_\textrm{M}=0.27$, $\Omega_\Lambda=0.73$, $\Omega_\textrm{K}=1-\Omega_\textrm{M}-\Omega_\Lambda$, $H_0=100\,$km$\,$s$^{-1}\,$Mpc$^{-1}$, and $P_\textrm{d}(\boldsymbol{k}_\perp,k_\parallel)$ is in units of K$^2$(Mpc/$h$)$^3$.

In summary, the delay spectrum, $\tilde{V}_b(\tau)$, is obtained from {\it visibilities} which are the basic data blocks measured by each antenna pair, using equations~\ref{eqn:obsvis} and \ref{eqn:delay-transform}. $\tilde{V}_b(\tau)$ captures all the effects of EoR H{\sc i} signal corruption caused by foregrounds and the instrument. At the same time, it is closely related to the sought power spectrum containing critical information about spatial scales. In using the delay spectrum technique, visibilities from different baselines have not been averaged together.

@@

We give a brief overview of some parameters of Fourier space which are generic to all experiments that use a similar approach. Figure~\ref{fig:fourier-space} illustrates the Fourier space in which the delay (and power) spectra of redshifted H{\sc i} observations are calculated. $|\boldsymbol{b}|$ and $k_\perp$, denoting spatial scales in the transverse direction (tangent plane to the celestial sphere), form the $x$--axis. $\tau$ and $k_\parallel$, denoting spatial scales along line--of--sight form the $y$--axis. Foreground emission maps to a wedge--shaped region in Fourier space, hereafter referred to as the {\it foreground wedge} \citep{dat10}, whose boundaries are determined by the antenna spacings and the light travel times across them. These boundaries, called horizon delay limits \citep{ved12,par12b}, are shown by solid lines.

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figure1.eps}
\caption{Fourier space in which delay (and power) spectra of EoR H{\sc i} signals are calculated. The $x$--axis is denoted by $|\boldsymbol{b}|$ (antenna spacing) or $k_\perp$ (transverse wavenumber). The $y$--axis denoted by $\tau$ (delay) or $k_\parallel$ (line--of--sight wavenumber). Here, $k_\perp$ and $k_\parallel$ are obtained for a frequency of 185~MHz. The dark shaded region is referred to as the {\it foreground wedge} where smooth spectrum foregrounds reside. Its boundaries (solid lines), given by light travel time for corresponding antenna spacings, are referred to as horizon delay limits. Narrow extensions of the {\it wedge} (white unshaded strips) are caused by convolution with the instrument's spectral transfer function. Regions excluding the {\it wedge} are expected to be relatively free of foreground contamination and are generally referred to as the {\it EoR window}. There are undesirable grating responses (dot--dashed lines) specific to the MWA. Hence, we conservatively identify a restricted region of high EoR sensitivity (medium shade) and refer to it as the MWA {\it EoR window}. \label{fig:fourier-space}}
\end{figure}

The spectral transfer function of the instrument convolves the {\it foreground wedge} and stretches it further (unshaded narrow strips bounded by solid and dashed lines) along $\tau$--axis \citep{par12b,thy13}. The width of this narrow strip is inversely proportional to the operating bandwidth. The region of Fourier space excluding the {\it foreground wedge} and the narrow strips is the so--called {\it EoR window}, shown in light and medium shades. In the context of EoR studies in Fourier space, the H{\sc i} power spectrum from the EoR is expected to decrease rapidly with $|\boldsymbol{k}|$. Hence, the brightest EoR signal will be observed on the shortest baselines and smallest delays. Thus the regions of interest for EoR studies relying on {\it avoidance} strategy are just beyond the horizon delay limits (dashed lines) on short baselines, marked as regions of {\it maximal EoR sensitivity}.

In the specific case of the MWA, which has a passband constructed using coarse channels, there are period grating responses resulting in repetitions of the {\it foreground wedge} at multiples of 0.78~$\mu$s. Thus the MWA {\it EoR window} lies outside the dashed lines but inside the first grating response ($|\tau| \lesssim 0.78\,\mu$s, dot--dashed lines) and is shown in medium shade.

@@

We obtain the delay spectrum of visibilities by taking the delay transform of each baseline's spectrum (equation~\ref{eqn:delay-transform}) choosing $W(f)$ to be a {\it Blackman--Harris} window function. The sky spectrum is multiplied in the instrument by the instrumental passband and flagging of frequency channels possibly corrupted by radio frequency interference, which together are represented by the weights, $W_\textrm{i}(f)$. In delay--space, these weights translate into a convolution by a point spread function (PSF). We deconvolve this PSF using a one dimensional CLEAN algorithm \citep{tay99} as described for the delay axis \citep{par09,par12b} to obtain the final delay spectra. The CLEAN procedure iteratively finds and subtracts peak values convolved by the Fourier transform of the weights. We limit the selection of peaks to modes inside the horizon delay limit, corresponding to smooth spectrum objects in the visible sky hemisphere.

\section{Wide--Field Measurements}\label{sec:wide-field}

With $\hat{\boldsymbol{s}}\equiv (l,m,n)$, equation~\ref{eqn:obsvis} can be written as \citep{tay99,tho01}:
\begin{align}\label{eqn:obsvis-lmn}
  V_b(f) &= \iint\limits_\textrm{sky} \frac{A(\hat{\boldsymbol{s}},f)\,I(\hat{\boldsymbol{s}},f)}{\sqrt{1-l^2-m^2}}\,W_\textrm{i}(f)\,e^{-i2\pi f\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}}\,\dif l\,\dif m, 
\end{align}
where, $l$, $m$, and $n$ denote the direction cosines toward East, North, and zenith respectively, with $n\equiv\sqrt{1-l^2-m^2}$, and:
\begin{align}\label{eqn:solid-angle}
  \dif\Omega = \frac{\dif l\,\dif m}{\sqrt{1-l^2-m^2}}.
\end{align}

When the synthesized field is small, where $A(\hat{\boldsymbol{s}},f)$ or $I(\hat{\boldsymbol{s}},f)$ is significant only for $|l| \ll 1$ and $|m| \ll 1$, equation~\ref{eqn:obsvis-lmn} reduces to a simple two--dimensional Fourier transform \citep{tay99,tho01} between the apparent sky brightness and measured visibilities. It is in this context that radio interferometers are understood to be sensitive only to fluctuations and not to a uniform sky brightness distribution.

In a wide--field measurement, neither $A(\hat{\boldsymbol{s}},f)$ nor $I(\hat{\boldsymbol{s}},f)$, in general, is negligible anywhere in the visible hemisphere. The solid angle per pixel on the sky in direction cosine coordinates changes significantly with direction (equation~\ref{eqn:solid-angle}), increasing steeply towards the horizon. Hence, the approximations in the narrow--field scenario do not apply. For example, even if $A(\hat{\boldsymbol{s}},f)$ and $I(\hat{\boldsymbol{s}},f)$ are held constant across the visible hemisphere, the amplitude of the integrand in equation~\ref{eqn:obsvis-lmn} is still dependent on direction. Therefore, in a significant departure from a narrow--field measurement, the wide--field visibility from a uniform brightness distribution on a non--zero antenna spacing is not zero.

Figure~\ref{fig:usm} shows the wide--field delay power spectrum of a uniformly illuminated sky with no spectral variation as measured by antenna elements arranged identical to that in the MWA antenna array layout \citep{bea12} with a uniform power pattern across the sky and a bandwidth of 30.72~MHz centered around 185~MHz (refer to \S\ref{sec:instrument_model} for a detailed description of the instrument model). Notice the steep rise in power towards the horizon limits. These wide--field effects are prevalent on all antenna spacings, including the longest ones used in this study.

>>>>

\title{Foregrounds in Wide-Field Redshifted 21~cm Power Spectra}

@@

Detection of 21~cm emission of H~{\sc i} from the epoch of reionization, at redshifts $z>6$, is limited primarily by foreground emission. We investigate the signatures of wide-field measurements and an all-sky foreground model using the delay spectrum technique that maps the measurements to foreground object locations through signal delays between antenna pairs. We demonstrate interferometric measurements are inherently sensitive to all scales, including the largest angular scales, owing to the nature of wide-field measurements. These wide-field effects are generic to all observations but antenna shapes impact their amplitudes substantially. A dish-shaped antenna yields the most desirable features from a foreground contamination viewpoint, relative to a dipole or a phased array. Comparing data from recent Murchison Widefield Array observations, we demonstrate that the foreground signatures that have the largest impact on the H~{\sc i} signal arise from power received far away from the primary field of view. We identify diffuse emission near the horizon as a significant contributing factor, even on wide antenna spacings that usually represent structures on small scales. For signals entering through the primary field of view, compact emission dominates the foreground contamination. These two mechanisms imprint a characteristic {\it pitchfork} signature on the ``foreground wedge'' in Fourier delay space. Based on these results, we propose that selective down-weighting of data based on antenna spacing and time can mitigate foreground contamination substantially by a factor $\sim 100$ with negligible loss of sensitivity.

@@

At the end of the recombination epoch, the Universe was completely neutral. This period, referred to as the {\it Dark Ages} in the Universe's history, is characterized by the localized accumulation of matter under the influence of gravity. And it ended with the formation of the first stars and galaxies which started emitting ultra-violet and X-ray radiation, thereby reionizing the neutral medium in their surroundings. This commenced the epoch of reionization (EoR) -- a period of nonlinear growth of matter density perturbations and astrophysical evolution.

Observing redshifted 21~cm radiation generated by the spin flip transition of H~{\sc i} has been identified as a direct probe of the EoR \citep{sun72,sco90,mad97,toz00,ili02}. Detecting this signal has recently emerged as a very promising experiment to fill the gaps in our understanding of the Universe's history.

Sensitive instruments such as the Square Kilometre Array (SKA) are required for direct observation and tomography of redshifted H~{\sc i}. Numerous pathfinders and precursors to the SKA such as the Murchison Widefield Array \citep[MWA;][]{lon09,tin13,bow13}, the Low Frequency Array \citep[LOFAR;][]{van13}, and the Precision Array for Probing the Epoch of Reionization \citep[PAPER;][]{par10} have become operational with enough sensitivity for a statistical detection of the EoR H~{\sc i} power spectrum \citep{bow06,par12a,bea13,dil13,thy13,pob14}. The Hydrogen Epoch of Reionization Array\footnote{\url{http://reionization.org/}} (HERA) is currently under construction using new insights gained with the MWA and PAPER.

A key challenge in the statistical detection of the redshifted H~{\sc i} 21~cm signal, via the spatial power spectrum of temperature fluctuations, arises from the contamination by Galactic and extragalactic foregrounds \citep[see, e.g.,][]{dim02,zal04,fur06,ali08,ber09,ber10,gho12}. \citet{mor04} show that the inherent isotropy and symmetry of the EoR signal in frequency and spatial wavenumber ($k$) space make it distinguishable from sources of contamination which are isolated to certain $k$ modes by virtue of their inherent spectral smoothness \citep{mor06,bow09,liu11,par12b,dil13,pob13}. Since this contamination is expected to be several orders of magnitude stronger than the underlying EoR H~{\sc i} signal, it is critical to characterize foregrounds precisely in order to reduce their impact on EoR H~{\sc i} power spectrum detection sensitivity.

Considerable effort is being made toward understanding the $k$-space behavior of foreground signatures in the observed power spectrum and formulating robust estimators of the true power spectrum \citep{bow09,liu09,dat10,liu11,mor12,tro12,pob13,thy13,dil14,liu14a,liu14b}. A model that provides a generic explanation for the observed foreground power spectrum has emerged, whereby the wide-field (and chromatic) response of the instrument causes the power in smooth spectrum foregrounds to occupy higher $k$-modes into the so-called ``wedge''. The conservative foreground strategy, referred to as {\it avoidance}, that has developed alongside this work is to discard $k$-modes which could be contaminated \citep[e.g.,][]{par14}. The more aggressive alternative is to subtract a sky model and regain access to modes that would be discarded by {\it avoidance}. In both cases, which parts of the sky are most critical to either {\it avoid} or {\it subtract} has remained largely uncertain. Here, we focus primarily on extending the {\it avoidance} strategy by identifying foreground components at greatest risk to ``leak'' from foreground modes to EoR modes and proposing a scheme for down-weighting these components.

Foregrounds with intrinsic deviations from spectral smoothness, instruments with high chromaticity, polarization leakage, calibration errors, or approximations in power spectrum analyses can contaminate the true EoR H~{\sc i} power spectrum. Here we use existing catalogs and a high fidelity instrumental model to capture both foreground and instrumental chromaticity. To decouple these effects from possible analysis effects, such as those pointed out by \citet{haz13}, we compute power spectra using a {\it per-baseline} approach of \citet{par12b}. This approximates the power spectrum as the inverse Fourier transform of the spectra generated by the instrument's correlator.

In \S\ref{sec:delay-spectrum} we provide an overview of the delay spectrum technique. We investigate signatures generic to all wide-field measurements of EoR power spectra in \S\ref{sec:wide-field}. In \S\ref{sec:sim}, we present the foreground model and a variety of instrument models to rank antenna shapes based on foreground contamination. In \S\ref{sec:MWA}, we describe the MWA setup, summarize the observing parameters, and present the resulting data. Simulations using these observing parameters are compared with the data and analyzed for foreground signatures. We report two important findings: foregrounds that most severely obscure the redshifted 21~cm power spectrum are not caused by emission in the central field of view, but rather by bright objects from near the horizon; and, diffuse Galactic emission plays a significant role hitherto unpredicted. In \S\ref{sec:fg-grading}, we offer an initial description of a more precise foreground avoidance technique that minimizes foreground contamination using prior knowledge of the sky to down-weight adversely contaminated baselines. We present a summary of our work and findings in \S\ref{sec:summary}.

@@

Interferometer array data known as {\it visibilities}, $V_b(f)$, represent correlations between time-series of electric fields measured by different antenna pairs with separation vectors $\boldsymbol{b}$ and then Fourier transformed along the time axis to obtain a spectrum along the frequency ($f$) axis. If $I(\hat{\boldsymbol{s}},f)$ and $A(\hat{\boldsymbol{s}},f)$ are the sky brightness and antenna's directional power pattern, respectively, at different frequencies as a function of direction on the sky denoted by the unit vector ($\hat{\boldsymbol{s}}$), and $W_\textrm{i}(f)$ denotes instrumental bandpass weights, then $V_b(f)$ can be written as:
\begin{align}\label{eqn:obsvis}
  V_b(f) &= \iint\limits_\textrm{sky} A(\hat{\boldsymbol{s}},f)\,I(\hat{\boldsymbol{s}},f)\,W_\textrm{i}(f)\,e^{-i2\pi f\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}}\,\dif\Omega,
\end{align}
where, $c$ is the speed of light, and $\dif\Omega$ is the solid angle element to which $\hat{\boldsymbol{s}}$ is the unit normal vector. This equation is valid in general, including wide-field measurements, and is a slight adaptation from \citet{van34}, \citet{zer38}, and \citet{tho01}.

The {\it delay spectrum}, $\tilde{V}_b(\tau)$, is defined as the inverse Fourier transform of $V_b(f)$ along the frequency coordinate:
\begin{align}\label{eqn:delay-transform}
  \tilde{V}_b(\tau) &\equiv \int V_b(f)\,W(f)\,e^{i2\pi f\tau}\,\dif f,
\end{align}
where, $W(f)$ is a spectral weighting function which can be chosen to control the quality of the delay spectrum \citep{ved12,thy13}, and $\tau$ represents the signal delay between antenna pairs:
\begin{equation}\label{eqn:delay}
  \tau = \frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}.
\end{equation}
The delay transform conventions used in this paper are described in appendix~\ref{sec:delay-conventions}. $\tilde{V}_b(\tau)$ is expressed in observer's units of Jy~Hz.

The delay spectrum has a close resemblance to cosmological H~{\sc i} spatial power spectrum. Appendix~\ref{sec:EoR-power-spectrum} gives an overview of the similarities and differences between the two. Foregrounds can be described in either framework. For our study, we find the delay spectrum approach to be simple and yet extremely useful.

In order to express a quantity derived from $\tilde{V}_b(\tau)$ whose units are the same as that of the cosmological H~{\sc i} power spectrum, we define the delay power spectrum:
\begin{align}\label{eqn:delay-power-spectrum}
  P_\textrm{d}(\boldsymbol{k}_\perp,k_\parallel) &\equiv |\tilde{V}_b(\tau)|^2\left(\frac{A_\textrm{e}}{\lambda^2\Delta B}\right)\left(\frac{D^2\Delta D}{\Delta B}\right)\left(\frac{\lambda^2}{2k_\textrm{B}}\right)^2,
\end{align}
with
\begin{align}
  \boldsymbol{k}_\perp &\equiv \frac{2\pi(\frac{\boldsymbol{b}}{\lambda})}{D}, \\
  k_\parallel &\equiv \frac{2\pi\tau\,f_{21}H_0\,E(z)}{c(1+z)^2}, 
\end{align}
where, $A_\textrm{e}$ is the effective area of the antenna, $\Delta B$ is the bandwidth, $\lambda$ is the wavelength of the band center, $k_\textrm{B}$ is the Boltzmann constant, $f_{21}$ is the rest frame frequency of the 21~cm spin flip transition of H~{\sc i}, $z$ is the redshift, $D\equiv D(z)$ is the transverse comoving distance, $\Delta D$ is the comoving depth along the line of sight corresponding to $\Delta B$, and $h$, $H_0$ and $E(z)\equiv [\Omega_\textrm{M}(1+z)^3+\Omega_\textrm{k}(1+z)^2+\Omega_\Lambda]^{1/2}$ are standard terms in cosmology. Throughout the paper, we use $\Omega_\textrm{M}=0.27$, $\Omega_\Lambda=0.73$, $\Omega_\textrm{K}=1-\Omega_\textrm{M}-\Omega_\Lambda$, $H_0=100\,$km$\,$s$^{-1}\,$Mpc$^{-1}$, and $P_\textrm{d}(\boldsymbol{k}_\perp,k_\parallel)$ is in units of K$^2$(Mpc/$h$)$^3$.

In summary, the delay spectrum, $\tilde{V}_b(\tau)$, is obtained from {\it visibilities} which are the basic data blocks measured by each antenna pair, using equations~\ref{eqn:obsvis} and \ref{eqn:delay-transform}. $\tilde{V}_b(\tau)$ captures all the effects of EoR H~{\sc i} signal corruption caused by foregrounds and the instrument. At the same time, it is closely related to the sought power spectrum containing critical information about spatial scales. In using the delay spectrum technique, visibilities from different baselines have not been averaged together.

@@

We give a brief overview of some parameters of Fourier space which are generic to all experiments that use a similar approach. Figure~\ref{fig:fourier-space} illustrates the Fourier space in which the delay (and power) spectra of redshifted H~{\sc i} observations are calculated. $|\boldsymbol{b}|$ and $k_\perp$, denoting spatial scales in the transverse direction (tangent plane to the celestial sphere), form the $x$-axis. $\tau$ and $k_\parallel$, denoting spatial scales along line of sight form the $y$-axis. Foreground emission maps to a wedge-shaped region in Fourier space, hereafter referred to as the {\it foreground wedge} \citep{dat10}, whose boundaries are determined by the antenna spacings and the light travel times across them. These boundaries, called horizon delay limits \citep{ved12,par12b}, are shown by solid lines.

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figure1.eps}
\caption{Fourier space in which delay (and power) spectra of EoR H~{\sc i} signals are calculated. The $x$-axis is denoted by $|\boldsymbol{b}|$ (antenna spacing) or $k_\perp$ (transverse wavenumber). The $y$-axis denoted by $\tau$ (delay) or $k_\parallel$ (line of sight wavenumber). Here, $k_\perp$ and $k_\parallel$ are obtained for a frequency of 185~MHz. The dark shaded region is referred to as the {\it foreground wedge} where smooth spectrum foregrounds reside. Its boundaries (solid lines), given by light travel time for corresponding antenna spacings, are referred to as horizon delay limits. Narrow extensions of the {\it wedge} (white unshaded strips) are caused by convolution with the instrument's spectral transfer function. Regions excluding the {\it wedge} are expected to be relatively free of foreground contamination and are generally referred to as the {\it EoR window}. There are undesirable grating responses (dotted-dashed lines) specific to the MWA. Hence, we conservatively identify a restricted region of high EoR sensitivity (medium shade) and refer to it as the MWA {\it EoR window}. \label{fig:fourier-space}}
\end{figure}

The spectral transfer function of the instrument convolves the {\it foreground wedge} and stretches it further (unshaded narrow strips bounded by solid and dashed lines) along $\tau$-axis \citep{par12b,thy13}. The width of this narrow strip is inversely proportional to the operating bandwidth. The region of Fourier space excluding the {\it foreground wedge} and the narrow strips is the so-called {\it EoR window}, shown in light and medium shades. In the context of EoR studies in Fourier space, the H~{\sc i} power spectrum from the EoR is expected to decrease rapidly with $|\boldsymbol{k}|$. Hence, the brightest EoR signal will be observed on the shortest baselines and smallest delays. Thus the regions of interest for EoR studies relying on {\it avoidance} strategy are just beyond the horizon delay limits (dashed lines) on short baselines, marked as regions of {\it maximal EoR sensitivity}.

In the specific case of the MWA, which has a passband constructed using coarse channels, there are period grating responses resulting in repetitions of the {\it foreground wedge} at multiples of 0.78~$\mu$s. Thus the MWA {\it EoR window} lies outside the dashed lines but inside the first grating response ($|\tau| \lesssim 0.78\,\mu$s, dotted-dashed lines) and is shown in medium shade.

@@

We obtain the delay spectrum of visibilities by taking the delay transform of each baseline's spectrum (equation~\ref{eqn:delay-transform}) choosing $W(f)$ to be a {\it Blackman-Harris} window function. The sky spectrum is multiplied in the instrument by the instrumental passband and flagging of frequency channels possibly corrupted by radio frequency interference, which together are represented by the weights, $W_\textrm{i}(f)$. In delay-space, these weights translate into a convolution by a point spread function (PSF). We deconvolve this PSF using a one dimensional CLEAN algorithm \citep{tay99} as described for the delay axis \citep{par09,par12b} to obtain the final delay spectra. The CLEAN procedure iteratively finds and subtracts peak values convolved by the Fourier transform of the weights. We limit the selection of peaks to modes inside the horizon delay limit, corresponding to smooth spectrum objects in the visible sky hemisphere.

\section{Wide-Field Measurements}\label{sec:wide-field}

With $\hat{\boldsymbol{s}}\equiv (l,m,n)$, equation~\ref{eqn:obsvis} can be written as \citep{tay99,tho01}:
\begin{align}\label{eqn:obsvis-lmn}
  V_b(f) &= \iint\limits_\textrm{sky} \frac{A(\hat{\boldsymbol{s}},f)\,I(\hat{\boldsymbol{s}},f)}{\sqrt{1-l^2-m^2}}\,W_\textrm{i}(f)\,e^{-i2\pi f\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}}\,\dif l\,\dif m, 
\end{align}
where, $l$, $m$, and $n$ denote the direction cosines toward east, north, and zenith respectively, with $n\equiv\sqrt{1-l^2-m^2}$, and:
\begin{align}\label{eqn:solid-angle}
  \dif\Omega = \frac{\dif l\,\dif m}{\sqrt{1-l^2-m^2}}.
\end{align}

When the synthesized field is small, where $A(\hat{\boldsymbol{s}},f)$ or $I(\hat{\boldsymbol{s}},f)$ is significant only for $|l| \ll 1$ and $|m| \ll 1$, equation~\ref{eqn:obsvis-lmn} reduces to a simple two-dimensional Fourier transform \citep{tay99,tho01} between the apparent sky brightness and measured visibilities. It is in this context that radio interferometers are understood to be sensitive only to fluctuations and not to a uniform sky brightness distribution.

In a wide-field measurement, neither $A(\hat{\boldsymbol{s}},f)$ nor $I(\hat{\boldsymbol{s}},f)$, in general, is negligible anywhere in the visible hemisphere. The solid angle per pixel on the sky in direction cosine coordinates changes significantly with direction (equation~\ref{eqn:solid-angle}), increasing steeply toward the horizon. Hence, the approximations in the narrow-field scenario do not apply. For example, even if $A(\hat{\boldsymbol{s}},f)$ and $I(\hat{\boldsymbol{s}},f)$ are held constant across the visible hemisphere, the amplitude of the integrand in equation~\ref{eqn:obsvis-lmn} is still dependent on direction. Therefore, in a significant departure from a narrow-field measurement, the wide-field visibility from a uniform brightness distribution on a non-zero antenna spacing is not zero.

Figure~\ref{fig:usm} shows the wide-field delay power spectrum of a uniformly illuminated sky with no spectral variation as measured by antenna elements arranged identical to that in the MWA antenna array layout \citep{bea12} with a uniform power pattern across the sky and a bandwidth of 30.72~MHz centered around 185~MHz (refer to \S\ref{sec:instrument_model} for a detailed description of the instrument model). Notice the steep rise in power toward the horizon limits. These wide-field effects are prevalent on all antenna spacings, including the longest ones used in this study.
